A comparison of two second order traffic flow models

### 2015-03-13 FirstDraftMaterials/Draft.tex
Modified conclusion

At first glance, this seems like a reasonable model.  One fundamental difference between cars and fluids causes this model to predict nonphysical situations.  As pointed out by Daganzo, 	``a fluid particle responds to stimuli from the front and from behind, but a car is an anisotropic particle that mostly responds to frontal stimuli" \cite{Daganzo1995}.  If we consider the eigenvalues of the linearized system,

\[ \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_t + \begin{pmatrix} v & \rho \\[1ex] \dfrac{p'(\rho)}{\rho} & v \end{pmatrix} \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_x = \begin{pmatrix} 0 \\[1ex] 0 \end{pmatrix}, \]

@@

% AR Model: Theory
\subsection{Theory}
To illustrate the PW model's problematic representation of cars as a fluid, Aw and Rascle \cite{AwRascle2000} consider the following situation.  Assume you are in your car, traveling with velocity $v$.  If the density ahead of you is increasing with respect to $x$ but decreasing with respect to $x - vt$, would you speed up or slow down?  Although the cars ahead are more densely spaced, they are traveling faster than you, so you would most likely accelerate.  However, since the density ahead is increasing with respect to $x$, the PW model predicts that you would slow down.

@@

we see that one issues is resolved.  In this model, the largest eigenvalue $\lambda_2 = v$.  Therefore no characteristic speed is greater than the velocity, so that drivers are no longer influenced by conditions behind them. To validate their model, Aw and Rascle claim that their improved model satisfies the following important conditions:
\begin{enumerate}
\item The system is hyperbolic.
\item When solving the Riemann problem with bounded nonnegative data $(\rho, v)$, the density and velocity remain nonnegative and bounded from above.
\item When solving the Riemann problem, no waves connecting any state to its left (behind it) have a propagation speed greater than the velocity $v$.
\item The solution to the Riemann problem agrees with qualitative properties that drivers actually experience: braking produces shock waves and acceleration produces rarefaction waves.
\end{enumerate}

@@

\section{Conclusion}
By studying the examples of the LWR, PW and AR models for traffic flow, we see that higher order models are not always more accurate than first order models.  Additionally, the assumptions made in developing higher order relations must conform to the phenomena they are designed to model.  For example, using equations for isotropic fluids to model anisotropic cars in the PW model produces nonphysical results.  The AR model corrects this problem by using a convective derivative rather than a space derivative for the anticipation factor.  Therefore, whenever we introduce higher order models, we should follow the example of Aw and Rascle by validating the model's results in a systematic way.

>>>>

At first glance, this seems like a reasonable model.  However, one fundamental difference between cars and fluids causes this model to predict nonphysical situations.  As pointed out by Daganzo, 	``a fluid particle responds to stimuli from the front and from behind, but a car is an anisotropic particle that mostly responds to frontal stimuli" \cite{Daganzo1995}.  If we consider the eigenvalues of the linearized system,

\[ \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_t + \begin{pmatrix} v & \rho \\[1ex] \dfrac{p'(\rho)}{\rho} & v \end{pmatrix} \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_x = \begin{pmatrix} 0 \\[1ex] \dfrac{V(\rho) - v}{\tau} + \mu v_{xx} \end{pmatrix}, \]

@@

% AR Model: Theory
\subsection{Theory}
To illustrate the PW model's problematic representation of cars as a fluid, Aw and Rascle \cite{AwRascle2000} consider the following situation.  Assume you are in your car, traveling with velocity $v$.  If the density ahead of you is increasing with respect to $x$ but decreasing with respect to $x - vt$, would you speed up or slow down?  Although the cars ahead are more densely packed, they are traveling faster than you, so you would most likely accelerate.  However, since the density ahead is increasing with respect to $x$, the PW model predicts that you would slow down.

@@

we see that one issues is resolved.  In this model, the largest eigenvalue is $\lambda_2 = v$.  Therefore no characteristic speed is greater than the velocity, so that drivers are no longer influenced by conditions behind them. To validate their model, Aw and Rascle claim that their improved model satisfies the following important conditions:
\begin{enumerate}
\item The system is hyperbolic.
\item When solving the Riemann problem with bounded nonnegative data $(\rho, v)$, the density and velocity remain nonnegative and bounded from above.
\item When solving the Riemann problem, no waves connecting any state to its left (behind it) have a propagation speed greater than the velocity $v$.
\item The solution to the Riemann problem agrees with qualitative properties that drivers actually experience: braking produces shock waves and acceleration produces rarefaction waves.
\end{enumerate}

@@

\section{Conclusion}
By studying the examples of the LWR, PW and AR models for traffic flow, we see that higher order models are not always more accurate than first order models.  Additionally, the assumptions made in developing higher order relations must conform to the phenomena they are designed to model.  For example, using equations for isotropic fluids to model anisotropic cars in the PW model produces nonphysical results.  The AR model corrects this problem by using a convective derivative rather than a space derivative for the anticipation factor.  Furthermore, the authors verified that their model behaved reasonably by checking it against a list of important conditions.  Therefore, whenever we introduce higher order models, we should follow the example of Aw and Rascle by validating the model's results in a systematic way.
